Informing policy via dynamic models: Eliminating cholera in Haiti

### 2022-10-18 ms.Rnw
Minor updates, code cleaning

The latent state is described as $\bm{X}^{(3)}(t) = \big(S_{u\vaccCounter}(t),\allowbreak I_{u\vaccCounter}(t),\allowbreak A_{u\vaccCounter}(t),\allowbreak R_{u\vaccCounter k}(t),\allowbreak W_u(t),\allowbreak u \in \seq{0}{U},\allowbreak \vaccCounter \in \seq{0}{4},\allowbreak k \in \seq{1}{3}\big)$.
Here, $\vaccCounter=0$ corresponds to unvaccinated, $\vaccCounter=2j-1$ corresponds to a single dose on the $j$th vaccination campaign in unit $u$ and $\vaccCounter=2j$ corresponds to receiving two doses on the $j$th vaccination campaign.
$k\in\seq{1}{3}$ models non-exponential duration in the recovered class before waning of immunity.
The force of infection is
\begin{equation}
\label{eq:model3:foi}
\lambda_u(t) = \Wbeta{_u} \frac{W_u(t)}{1+W_u(t)} + \transmission_u \sum_{{\varv}\neq u}\big(I_{\varv 0}(t)+ \asymptomRelativeInfect A_{\varv 0}(t)\big),
\end{equation}
and Per-capita transition rates are given in Equations~\ref{eq:model3:SI}--\ref{eq:model3:Decay}.
\begin{eqnarray}
\label{eq:model3:SI}
\mu_{S_{u\vaccCounter}I_{u\vaccCounter}} &=& \symptomFrac \,  \lambda_u \big(1-\eta_{u\vaccCounter}(t)\big) \, d\Gamma/dt,
\\[-3pt]
\label{eq:model3:SA}
\mu_{S_{u\vaccCounter}A_{u\vaccCounter}} &=& (1-\symptomFrac) \,  \lambda_u \big(1-\eta_{u\vaccCounter}(t)\big) \,  d\Gamma/dt,
\\[-3pt]
\label{eq:model3:IR}
\mu_{I_{u\vaccCounter}R_{u\vaccCounter 1}} &=& \mu_{A_{u\vaccCounter}R_{u\vaccCounter 1}} = \muIR,
\\[-3pt]
\label{eq:model3:IS}
\mu_{I_{u\vaccCounter}S_{u0}} &=& \muDeath + \choleraDeath,
\\[-3pt]
\label{eq:model3:AS}
\mu_{A_{u\vaccCounter}S_{u0}} &=& \muDeath,
\\[-3pt]
\label{eq:model3:RRnext}
\mu_{R_{u\vaccCounter 1}R_{u\vaccCounter 2}} &=& \mu_{R_{u\vaccCounter 2}R_{u\vaccCounter 3}} = 3\muRS,
\\[-3pt]
\label{eq:model3:RS}
\mu_{R_{u\vaccCounter k}S_{u0}} &=& \muDeath + 3\muRS \, \mathbf{1}_{\{k=3\}},
\\[-3pt]
\label{eq:model3:water}
\mu_{{\demography}W_u} &=& \big[1 + \seasAmplitude \big(J_u(t))^r \big] \mathrm{Den}_u \, \Wshed \big[ I_{u0}(t)+ \asymptomRelativeShed A_{u0}(t) \big],
\\[-3pt]
\label{eq:model3:Decay}
\mu_{W_u\demography} &=& \Wremoval.
\end{eqnarray}
As with Model~1, $d\Gamma_u(t)/dt$ is multiplicative Gamma-distributed white noise in \myeqref{eq:model3:SI} and \myeqref{eq:model3:SA}.
In \eqref{eq:model3:water}, $J_u(t)$ is a dimensionless measurement of precipitation that has been standardized by dividing the observed rainfall at time $t$ by the maximum recorded rainfall in department $u$ during the epidemic, and $\mathrm{Den}_u$ is the population density.
Demographic stochasticity is accounted for by modeling non-cholera related death rate $\muDeath$ in each compartment, along with an additional death rate $\choleraDeath$ in \myeqref{eq:model3:IS} to account for cholera induced deaths among infected individuals.
All deaths are balanced by births into the susceptible compartment in \myeqref{eq:model3:AS} and \myeqref{eq:model3:RS}, thereby maintaining constant population in each department.

@@

# Create vectors for the unit and shared parameters
unit_specific_names <- c("betaB", "foi_add", "Binit")
shared_param_names <- c(
  "mu_B", "XthetaA", "thetaI", "lambdaR", "r", "std_W",
  "epsilon", "k", "sigma"
)
est_param_names <- c(
  unit_specific_names, shared_param_names
)

@@

setrw1 <- function(x) {
  if (grepl('^r[[:digit:]]+$', x)) {
    return(0.01 / 2)
  } else if (grepl('^Binit[[:digit:]]+$', x)) {
    return(expression(ivp(0.02)))
  } else {
    return(0.01)
  }
}

setrw2 <- function(x) {
  if (grepl('^r[[:digit:]]+$', x)) {
    return(0.0025 / 2)
  } else if (grepl('^Binit[[:digit:]]+$', x)) {
    return(expression(ivp(0.005)))
  } else {
    return(0.0025)
  }
}

setrw3 <- function(x) {
  if (grepl('^r[[:digit:]]+$', x)) {
    return(0.001 / 2)
  } else if (grepl('^Binit[[:digit:]]+$', x)) {
    return(expression(ivp(0.002)))
  } else {
    return(0.001)
  }
}

# First Search RW
reg_rw_1.sd <- lapply(est_param_names_expanded, setrw1)
names(reg_rw_1.sd) <- est_param_names_expanded
chol_rw_1.sd <- do.call(rw.sd, reg_rw_1.sd)

# Second Search RW
reg_rw_2.sd <- lapply(est_param_names_expanded, setrw2)
names(reg_rw_2.sd) <- est_param_names_expanded
chol_rw_2.sd <- do.call(rw.sd, reg_rw_2.sd)

# Third Search RW
reg_rw_3.sd <- lapply(est_param_names_expanded, setrw3)
names(reg_rw_3.sd) <- est_param_names_expanded
chol_rw_3.sd <- do.call(rw.sd, reg_rw_3.sd)

# Search hyper-parameters for each run-level and search
if (RUN_LEVEL == 1){
  SEARCH1 = list(
    NBPF = 5,
    NP = 50,
    SPAT_REGRESSION = 0.5,
    NREPS = 6,
    NP_EVAL = 100,
    NREPS_EVAL = 6,
    RW_SD = chol_rw_1.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE
  )

  SEARCH2 = list(
    TOP_N = 2,
    NBPF = 5,
    NP = 50,
    SPAT_REGRESSION = 0.5,
    NREPS = 3,
    NP_EVAL = 100,
    NREPS_EVAL = 6,
    RW_SD = chol_rw_2.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE
  )

  SEARCH3 = list(
    TOP_N = 2,
    NBPF = 5,
    NP = 100,
    SPAT_REGRESSION = 0.5,
    NREPS = 3,
    NP_EVAL = 100,
    NREPS_EVAL = 6,
    RW_SD = chol_rw_3.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE
  )

  N_SEARCHES <- 2L
} else if (RUN_LEVEL == 2) {
  SEARCH1 = list(
    NBPF = 50,
    NP = 1000,
    SPAT_REGRESSION = 0.5,
    NREPS = 36,
    NP_EVAL = 1000,
    NREPS_EVAL = 36,
    RW_SD = chol_rw_1.sd,
    COOLING = 0.36,
    KEEP_TRACES = TRUE
  )

  SEARCH2 = list(
    TOP_N = 6,
    NBPF = 50,
    NP = 1000,
    SPAT_REGRESSION = 0.5,
    NREPS = 6,
    NP_EVAL = 1000,
    NREPS_EVAL = 36,
    RW_SD = chol_rw_2.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE
  )

  SEARCH3 = list(
    TOP_N = 6,
    NBPF = 50,
    NP = 1000,
    SPAT_REGRESSION = 0.5,
    NREPS = 6,
    NP_EVAL = 1000,
    NREPS_EVAL = 36,
    RW_SD = chol_rw_3.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE
  )

  N_SEARCHES <- 3L
} else if (RUN_LEVEL == 3) {

  SEARCH1 = list(
    NBPF = 100,
    NP = 10000,
    SPAT_REGRESSION = 0.5,
    NREPS = 72,
    NP_EVAL = 10000,
    NREPS_EVAL = 36,
    RW_SD = chol_rw_1.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE
  )

  SEARCH2 = list(
    TOP_N = 9,
    NBPF = 100,
    NP = 10000,
    SPAT_REGRESSION = 0.5,
    NREPS = 4,
    NP_EVAL = 10000,
    NREPS_EVAL = 36,
    RW_SD = chol_rw_2.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE
  )

  SEARCH3 = list(
    TOP_N = 6,
    NBPF = 100,
    NP = 10000,
    SPAT_REGRESSION = 0.5,
    NREPS = 6,
    NP_EVAL = 10000,
    NREPS_EVAL = 36,
    RW_SD = chol_rw_3.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE
  )

  N_SEARCHES <- 3L
}

if (RUN_LEVEL != 3 & file.exists("model3/run_level_3/haiti3_coupled_fit.rds")) {
  haiti3_fit <- readRDS("model3/run_level_3/haiti3_coupled_fit.rds")
} else {
  haiti3_fit <- bake(
    file = paste0("model3/", rl_dir, "haiti3_coupled_fit.rds"), {
      fit_coupled_haiti3(
        nsearches = N_SEARCHES,
        search1 = SEARCH1,
        search2 = SEARCH2,
        search3 = SEARCH3,
        ncores = cores
      )
    }
  )
}

@@

# Just a bit of tidying up: removing parameters that we won't use again.
rm(
  SEARCH1, SEARCH2, SEARCH3, N_SEARCHES, unit_specific_names,
  shared_param_names, est_param_names, est_param_names_expanded,
  setrw1, setrw2, setrw3, reg_rw_1.sd, reg_rw_2.sd, reg_rw_3.sd,
  chol_rw_1.sd, chol_rw_2.sd, chol_rw_3.sd
)
gc()

@@

p3_specific <- c("foi_add", "betaB", "Binit")
p3_shared <- c(
  "mu_B", "XthetaA", "thetaI", "lambdaR", "r", "std_W",
  "epsilon", "k", "sigma"
)

p3u <- matrix(nrow = length(p3_specific), ncol = 10)
rownames(p3u) <- p3_specific
p3s <- numeric(length(p3_shared))
names(p3s) <- p3_shared

for (p in p3_shared) {
  pat <- paste0("^", p, "[[:digit:]]+$")
  p3s[p] <-  mean(p3[grepl(pat, names(p3))])
}

for (p in p3_specific) {
  pat <- paste0("^", p, "[[:digit:]]+$")
  p3u[p, ] <- p3[grepl(pat, names(p3))]
}

rm(p3_specific, p3_shared, p, pat)

@@

h3_spat_nsim <- switch(RUN_LEVEL, 20, 100, 500)

# Check if run_level_3 results exit, and use them if they do.
if (RUN_LEVEL != 3 & file.exists("model3/run_level_3/SpatPompSims.rds")) {
  h3Spat_quants <- readRDS("model3/run_level_3/SpatPompSims.rds")
} else {
  h3Spat_quants <- bake(
    file = paste0('model3/', rl_dir, 'SpatPompSims.rds'), {

      # Simulate from the fitted version of Model 3
      h3Spat_sims <- simulate(
        h3Spat, nsim = h3_spat_nsim, format = 'data.frame'
      )

      # For each department and time-point, calculate percentiles of the
      # simulations
      h3Spat_sims %>%
        rename(dep = unitname) %>%
        group_by(dep, time) %>%
        summarise(
          q05 = quantile(cases, 0.025, na.rm = T),
          mean = mean(cases, na.rm = T),
          q50 = quantile(cases, 0.5, na.rm = T),
          q95 = quantile(cases, 0.975, na.rm = T)
        ) %>%
        ungroup() %>%
        mutate(
          date = lubridate::date_decimal(time)
        ) %>%
        # filter(
        #   date >= as.Date(lubridate::round_date(lubridate::date_decimal(h3Spat@t0), unit = 'day'))
        #   ) %>%
        mutate(date = as.Date(lubridate::round_date(date, unit = 'day')))
    },
    timing = FALSE
  )
}

@@

ggplot() +
  geom_line(data = h3Spat_quants, aes(x = date, y = q50 + 1), col = 'blue') +
  geom_line(data = dep_plot_df, aes(x = date, y = cases + 1), col = 'black') +
  geom_ribbon(  # plot confidence region.
    data = h3Spat_quants,
    aes(x = date, ymin = q05 + 1, ymax = q95 + 1),
    alpha = 0.4
  ) +
  facet_wrap(
    ~factor(dep, levels = plot_order), nrow = 2, labeller = dep_labeller
  ) +
  labs(y = 'Log Reported Cholera Cases') +
  scale_x_date(  # Format the date on the X-axis.
    date_labels = "'%y",
    breaks = seq.Date(
      from = as.Date("2011-01-01"),
      to = as.Date("2019-01-01"),
      by = '2 years'
    )
  ) +
  theme(axis.title.x = element_blank()) +
  scale_y_log10(  # Plot on log-scale
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  )
@
\caption{\label{fig:h3spatsims}
Simulations from initial conditions using the spatially coupled version of Model~3.
The black curve represents true case count, the blue line the median of 500 simulations from the model, and the gray ribbons representing $95\%$ confidence interval.
}
\end{figure}

<<Remove Model3 spatSims, cache=FALSE, include=FALSE, message=FALSE, echo=FALSE>>=
rm(h3Spat_quants, h3_spat_nsim)
gc()
@

@@

# TODO: Not needed after fitting Model 3.
NREPS_EVAL <- switch(RUN_LEVEL,   3,   8,    36)
NP_EVAL    <- switch(RUN_LEVEL, 100, 500, 10000)

@@

# model 2 AIC
mod2_n_params <- 16  # TODO: adjust Number of parameters fit in Model 2
mod2_aic <- 2 * mod2_n_params - 2 * mod2_ll

@@

# model 3 AIC
mod3_n_params <- 29
mod3_aic <- 2 * mod3_n_params - 2 * mod3_ll

@@

# TODO: The code below will not be needed once we update the fitting of model 3.
# Calculate likelihood via block-particle filter
if (RUN_LEVEL != 3 & file.exists("model3/run_level_3/SpatPomp_res.rds")) {
  mod3_spatPomp_res <- readRDS("model3/run_level_3/SpatPomp_res.rds")
} else{
  mod3_spatPomp_res <- bake(
    file = paste0('model3/', rl_dir, 'SpatPomp_res.rds'), {

@@

      # Perform block-particle filter
      h3_bpf <- foreach(i = 1:NREPS_EVAL, .combine = c) %dopar% {
        bpfilter(h3Spat, Np = NP_EVAL, block_size = 1,
                 params = haitipkg:::panelParms_toSpatParms3(p3))
      }

@@

mod3_spatPomp_n_params <- 29
mod3_spatPomp_aic <- 2 * mod3_spatPomp_n_params - 2 * mod3_spatPomp_res$ll
mod3_spatPomp_res$subset_aic <- 2 * mod3_spatPomp_n_params - 2 * mod3_spatPomp_res$subset_ll

@@

  & {\fixed $\muIR^{-1}=\Sexpr{myround(365/p3["gamma"],1)}$ }

@@

  & {\fixed $\muDeath^{-1} = \Sexpr{myround(1/p3s["mu"],1)}$}

>>>>

The latent state is described as $\bm{X}^{(3)}(t) = \big(S_{u\vaccCounter}(t),\allowbreak I_{u\vaccCounter}(t),\allowbreak A_{u\vaccCounter}(t),\allowbreak R_{u\vaccCounter k}(t),\allowbreak W_u(t),\allowbreak u \in \seq{0}{U},\allowbreak \vaccCounter \in \seq{0}{4},\allowbreak k \in \seq{1}{3}\big)$.
Here, $\vaccCounter=0$ corresponds to unvaccinated, $\vaccCounter=2j-1$ corresponds to a single dose on the $j$th vaccination campaign in unit $u$ and $\vaccCounter=2j$ corresponds to receiving two doses on the $j$th vaccination campaign.
$k\in\seq{1}{3}$ models non-exponential duration in the recovered class before waning of immunity.
The force of infection is
\begin{equation}
\label{eq:model3:foi}
\lambda_u(t) = \Wbeta{_u} \frac{W_u(t)}{1+W_u(t)} + \transmission_u \sum_{{\varv}\neq u}\big(I_{\varv 0}(t)+ \asymptomRelativeInfect A_{\varv 0}(t)\big),
\end{equation}
and Per-capita transition rates are given in Equations~\ref{eq:model3:SI}--\ref{eq:model3:Decay}.
\begin{eqnarray}
\label{eq:model3:SI}
\mu_{S_{u\vaccCounter}I_{u\vaccCounter}} &=& \symptomFrac \,  \lambda_u \big(1-\eta_{u\vaccCounter}(t)\big) \, d\Gamma/dt,
\\[-3pt]
\label{eq:model3:SA}
\mu_{S_{u\vaccCounter}A_{u\vaccCounter}} &=& (1-\symptomFrac) \,  \lambda_u \big(1-\eta_{u\vaccCounter}(t)\big) \,  d\Gamma/dt,
\\[-3pt]
\label{eq:model3:IR}
\mu_{I_{u\vaccCounter}R_{u\vaccCounter 1}} &=& \mu_{A_{u\vaccCounter}R_{u\vaccCounter 1}} = \muIR,
\\[-3pt]
\label{eq:model3:IS}
\mu_{I_{u\vaccCounter}S_{u0}} &=& \muDeath + \choleraDeath, \hspace{2mm} \mu_{A_{u\vaccCounter}S_{u0}} = \muDeath
\\[-3pt]
\label{eq:model3:RRnext}
\mu_{R_{u\vaccCounter 1}R_{u\vaccCounter 2}} &=& \mu_{R_{u\vaccCounter 2}R_{u\vaccCounter 3}} = 3\muRS,
\\[-3pt]
\label{eq:model3:RS}
\mu_{R_{u\vaccCounter k}S_{u0}} &=& \muDeath + 3\muRS \, \mathbf{1}_{\{k=3\}},
\\[-3pt]
\label{eq:model3:water}
\mu_{{\demography}W_u} &=& \big[1 + \seasAmplitude \big(J_u(t))^r \big] \mathrm{Den}_u \, \Wshed \big[ I_{u0}(t)+ \asymptomRelativeShed A_{u0}(t) \big],
\\[-3pt]
\label{eq:model3:Decay}
\mu_{W_u\demography} &=& \Wremoval.
\end{eqnarray}
As with Model~1, $d\Gamma_u(t)/dt$ is multiplicative Gamma-distributed white noise in \myeqref{eq:model3:SI} and \myeqref{eq:model3:SA}.
In \eqref{eq:model3:water}, $J_u(t)$ is a dimensionless measurement of precipitation that has been standardized by dividing the observed rainfall at time $t$ by the maximum recorded rainfall in department $u$ during the epidemic, and $\mathrm{Den}_u$ is the population density.
Demographic stochasticity is accounted for by modeling non-cholera related death rate $\muDeath$ in each compartment, along with an additional death rate $\choleraDeath$ in \myeqref{eq:model3:IS} to account for cholera induced deaths among infected individuals.
All deaths are balanced by births into the susceptible compartment in \myeqref{eq:model3:IS} and \myeqref{eq:model3:RS}, thereby maintaining constant population in each department.

@@

# Create vectors for the unit and shared parameters
unit_specific_names <- c("betaB", "foi_add", "Binit")
shared_param_names <- c(
  "mu_B", "XthetaA", "thetaI", "lambdaR", "r", "std_W",
  "epsilon", "k", "sigma", "rho", "gamma"
)

mod3_n_params <- 10 * length(unit_specific_names) + length(shared_param_names)

est_param_names <- c(
  unit_specific_names, shared_param_names
)

@@

set_rw_1 <- function(x) {
  if (grepl('^foi_add[[:digit:]]+$', x) || grepl('^betaB[[:digit:]]+$', x)) {
    return(0.01 / 2)
  } else if (grepl('^Binit[[:digit:]]+$', x)) {
    return(expression(ivp(0.025)))
  } else {
    return(0.01)
  }
}

set_rw_2 <- function(x) {
  if (grepl('^foi_add[[:digit:]]+$', x) || grepl('^betaB[[:digit:]]+$', x)) {
    return(0.003 / 2)
  } else if (grepl('^Binit[[:digit:]]+$', x)) {
    return(expression(ivp(0.0075)))
  } else {
    return(0.003)
  }
}

set_rw_3 <- function(x) {
  if (grepl('^foi_add[[:digit:]]+$', x) || grepl('^betaB[[:digit:]]+$', x)) {
    return(0.001 / 2)
  } else if (grepl('^Binit[[:digit:]]+$', x)) {
    return(expression(ivp(0.002)))
  } else {
    return(0.001)
  }
}

# First Search RW
reg_rw_1.sd <- lapply(est_param_names_expanded, set_rw_1)
names(reg_rw_1.sd) <- est_param_names_expanded
chol_rw_1.sd <- do.call(rw.sd, reg_rw_1.sd)

# Second Search RW
reg_rw_2.sd <- lapply(est_param_names_expanded, set_rw_2)
names(reg_rw_2.sd) <- est_param_names_expanded
chol_rw_2.sd <- do.call(rw.sd, reg_rw_2.sd)

# Third Search RW
reg_rw_3.sd <- lapply(est_param_names_expanded, set_rw_3)
names(reg_rw_3.sd) <- est_param_names_expanded
chol_rw_3.sd <- do.call(rw.sd, reg_rw_3.sd)

# Search hyper-parameters for each run-level and search
if (RUN_LEVEL == 1){
  SEARCH1 = list(
    NBPF = 5,
    NP = 50,
    SPAT_REGRESSION = 0.05,
    NREPS = 6,
    NP_EVAL = 100,
    NREPS_EVAL = 5,
    RW_SD = chol_rw_1.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE
  )

  SEARCH2 = list(
    TOP_N = 2,
    NBPF = 5,
    NP = 50,
    SPAT_REGRESSION = 0.05,
    NREPS = 3,
    NP_EVAL = 100,
    NREPS_EVAL = 5,
    RW_SD = chol_rw_2.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE
  )

  SEARCH3 = list(
    TOP_N = 2,
    NBPF = 5,
    NP = 100,
    SPAT_REGRESSION = 0.05,
    NREPS = 3,
    NP_EVAL = 100,
    NREPS_EVAL = 6,
    RW_SD = chol_rw_3.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE
  )

  N_SEARCHES <- 2L
} else if (RUN_LEVEL == 2) {
  SEARCH1 = list(
    NBPF = 20,
    NP = 500,
    SPAT_REGRESSION = 0.05,
    NREPS = 36,
    NP_EVAL = 1000,
    NREPS_EVAL = 12,
    RW_SD = chol_rw_1.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE
  )

  SEARCH2 = list(
    TOP_N = 6,
    NBPF = 20,
    NP = 500,
    SPAT_REGRESSION = 0.05,
    NREPS = 6,
    NP_EVAL = 1000,
    NREPS_EVAL = 12,
    RW_SD = chol_rw_2.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE
  )

  SEARCH3 = list(
    TOP_N = 6,
    NBPF = 20,
    NP = 500,
    SPAT_REGRESSION = 0.05,
    NREPS = 6,
    NP_EVAL = 1000,
    NREPS_EVAL = 12,
    RW_SD = chol_rw_3.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE
  )

  N_SEARCHES <- 1L
} else if (RUN_LEVEL == 3) {

  SEARCH1 = list(
    NBPF = 100,
    NP = 5000,
    SPAT_REGRESSION = 0.05,
    NREPS = 144,
    NP_EVAL = 5000,
    NREPS_EVAL = 18,
    RW_SD = chol_rw_1.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE
  )

  SEARCH2 = list(
    TOP_N = 12,
    NBPF = 100,
    NP = 5000,
    SPAT_REGRESSION = 0.05,
    NREPS = 6,
    NP_EVAL = 5000,
    NREPS_EVAL = 18,
    RW_SD = chol_rw_2.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE
  )

  SEARCH3 = list(
    TOP_N = 9,
    NBPF = 100,
    NP = 5000,
    SPAT_REGRESSION = 0.05,
    NREPS = 8,
    NP_EVAL = 5000,
    NREPS_EVAL = 18,
    RW_SD = chol_rw_3.sd,
    COOLING = 0.5,
    KEEP_TRACES = TRUE
  )

  N_SEARCHES <- 3L
} else {
  stop("Run level must be 1, 2 or 3.")
}

if (RUN_LEVEL != 3 & file.exists("model3/run_level_3/haiti3_coupled_fit.rds")) {
  haiti3_fit <- readRDS("model3/run_level_3/haiti3_coupled_fit.rds")
} else {
  haiti3_fit <- bake(
    file = paste0("model3/", rl_dir, "haiti3_coupled_fit.rds"), {
      fit_coupled_haiti3(
        nsearches = N_SEARCHES,
        search1 = SEARCH1,
        search2 = SEARCH2,
        search3 = SEARCH3,
        ncores = cores,
        search_rho = TRUE,
        search_gamma = TRUE
      )
    }
  )
}

@@

# Just a bit of tidying up: removing parameters that we won't use again.
rm(
  SEARCH1, SEARCH2, SEARCH3, N_SEARCHES,
  est_param_names, est_param_names_expanded,
  set_rw_1, set_rw_2, set_rw_3, reg_rw_1.sd, reg_rw_2.sd, reg_rw_3.sd,
  chol_rw_1.sd, chol_rw_2.sd, chol_rw_3.sd
)
gc()

@@

p3u <- matrix(nrow = length(unit_specific_names), ncol = 10)
rownames(p3u) <- unit_specific_names
p3s <- numeric(length(shared_param_names))
names(p3s) <- shared_param_names

for (p in shared_param_names) {
  pat <- paste0("^", p, "[[:digit:]]+$")
  p3s[p] <-  mean(p3[grepl(pat, names(p3))])
}

for (p in unit_specific_names) {
  pat <- paste0("^", p, "[[:digit:]]+$")
  p3u[p, ] <- p3[grepl(pat, names(p3))]
}

rm(unit_specific_names, shared_param_names, p, pat)

@@

h3_nsim <- switch(RUN_LEVEL, 20, 100, 500)

# Check if run_level_3 results exit, and use them if they do.
if (RUN_LEVEL != 3 & file.exists("model3/run_level_3/h3Sims.rds")) {
  h3_quants <- readRDS("model3/run_level_3/h3Sims.rds")
} else {
  h3_quants <- bake(
    file = paste0('model3/', rl_dir, 'h3Sims.rds'), {

      # Simulate from the fitted version of Model 3
      h3_sims <- simulate(
        h3Spat, nsim = h3_nsim, format = 'data.frame'
      )

      # For each department and time-point, calculate percentiles of the
      # simulations
      h3_sims %>%
        rename(dep = unitname) %>%
        group_by(dep, time) %>%
        summarise(
          q05 = quantile(cases, 0.025, na.rm = T),
          mean = mean(cases, na.rm = T),
          q50 = quantile(cases, 0.5, na.rm = T),
          q95 = quantile(cases, 0.975, na.rm = T)
        ) %>%
        ungroup() %>%
        mutate(
          date = lubridate::date_decimal(time)
        ) %>%
        mutate(date = as.Date(lubridate::round_date(date, unit = 'day')))
    },
    timing = FALSE
  )
}

@@

ggplot() +
  geom_line(data = h3_quants, aes(x = date, y = q50 + 1), col = 'blue') +
  geom_line(data = dep_plot_df, aes(x = date, y = cases + 1), col = 'black') +
  geom_ribbon(  # plot confidence region.
    data = h3_quants,
    aes(x = date, ymin = q05 + 1, ymax = q95 + 1),
    alpha = 0.4
  ) +
  facet_wrap(
    ~factor(dep, levels = plot_order), nrow = 2, labeller = dep_labeller
  ) +
  labs(y = 'Log Reported Cholera Cases') +
  scale_x_date(  # Format the date on the X-axis.
    date_labels = "'%y",
    breaks = seq.Date(
      from = as.Date("2011-01-01"),
      to = as.Date("2019-01-01"),
      by = '2 years'
    )
  ) +
  theme(axis.title.x = element_blank()) +
  scale_y_log10(  # Plot on log-scale
    labels = scales::trans_format("log10", scales::math_format(10^.x))
  )
@
\caption{\label{fig:h3spatsims}
Simulations from initial conditions using the spatially coupled version of Model~3.
The black curve represents true case count, the blue line the median of 500 simulations from the model, and the gray ribbons representing $95\%$ confidence interval.
}
\end{figure}

<<Remove Model3 spatSims, cache=FALSE, include=FALSE, message=FALSE, echo=FALSE>>=
rm(h3_quants, h3_nsim)
gc()
@

@@

NREPS_EVAL <- switch(RUN_LEVEL,   3,   8,    36)
NP_EVAL    <- switch(RUN_LEVEL, 100, 500,  5000)

@@

# model 2 AIC
mod2_n_params <- h2_fit$n_fit_params
mod2_aic <- 2 * mod2_n_params - 2 * mod2_ll

@@

# model 3 AIC
mod3_aic <- 2 * mod3_n_params - 2 * mod3_ll

@@

# Calculate likelihood via block-particle filter
if (RUN_LEVEL != 3 & file.exists("model3/run_level_3/mod3_eval_res.rds")) {
  mod3_eval_res <- readRDS("model3/run_level_3/mod3_eval_res.rds")
} else{
  mod3_eval_res <- bake(
    file = paste0('model3/', rl_dir, 'mod3_eval_res.rds'), {

@@

      # Perform block-particle filter
      h3_bpf <- foreach(i = 1:NREPS_EVAL, .combine = c) %dopar% {
        bpfilter(h3Spat, Np = NP_EVAL, block_size = 1)
      }

@@

mod3_aic <- 2 * mod3_n_params - 2 * mod3_eval_res$ll
mod3_eval_res$subset_aic <- 2 * mod3_n_params - 2 * mod3_eval_res$subset_ll

@@

  & {\fixed $\muIR^{-1}=\Sexpr{myround(365/p3s["gamma"],1)}$ }

@@

  & {\fixed $\muDeath^{-1} = \Sexpr{myround(1/p3["mu1"],1)}$}
